---
title: "R Style, Programming, and Debugging"
output:
  revealjs::revealjs_presentation:
    theme: white
    center: true
    transition: none
    incremental: true
---

```{r setup, include=FALSE}
library(knitr)
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
```
# R Programming Style

## R Style Guide
While there is not universal agreement on programming style, there are two good examples:


1. Hadley Wickham's Style Guide:
[http://adv-r.had.co.nz/Style.html](http://adv-r.had.co.nz/Style.html)

2. Google R Style Guide: [https://google.github.io/styleguide/Rguide.xml](https://google.github.io/styleguide/Rguide.xml), which presents different options from Wickham's guide.

## Notation and Naming

- **File Names:** File names should end in .R and be meaningful.
    - Good: predict_ad_revenue.R
    - Bad: foo.R
- **Identifiers:** Don't use underscores or hypens in identifiers.
    - The preferred form for variable names is all lower case letters with words separated with dots (`variable.name`), but `VariableName` is also accepted.
    - Function names begin with capital letters and include no dots (`FunctionName`)

## Syntax
- **Spacing:** 
    - Place spaces around all operators (`==, +, ...`)
    - Do not place a space before a comma, but always place one after a comma.
    - Place a space before left parenthesis, except in a function call.
- **Assignment:**
    - Use `<-` not `=` for assignment.

## Operators in R
- Most mathematical operators are self explanatory, but here are a few more important operators.
    - `==` will test for equality. For example to determine if pi equals three, this can be evaluated with `pi == 3` in R and will return `r pi == 3 `. Note this operator returns a logical value.
    - `&` is the AND operator, so `TRUE & FALSE` will return `r TRUE & FALSE`.
    - `|` is the OR operator, so `TRUE | FALSE` will return `r TRUE | FALSE`.
    - `!` is the NOT operator, so `! TRUE` will return `r  ! TRUE`.
    - `^` permits power terms, so `4 ^ 2` returns `r 4^2` and `4 ^ .5` returns `r 4 ^ .5`.


## Exercise: Order of operations
Note that order of operations is important in writing R code.
```{r opt,eval=F}
4 - 2 ^ 2
(4 - 2) ^ 2
5 * 2 - 3 ^ 2
pi == 3
! TRUE & pi == 3
! (TRUE | FALSE) 
```

Evaluate all  expressions. Note `!` is R's not operator.

## Solution: Order of operations
The results of the R code are:
```{r opt2, }
4 - 2 ^ 2
(4 - 2) ^ 2
5 * 2 - 3 ^ 2
```

## Solution: Order of operations
The results of the R code are:
```{r }
pi == 3
! TRUE & pi == 3
! (TRUE | FALSE) 
```


## Organization: Layout of .R
- **General Layout:**
The general layout of an R script should follow as:
  1. Author Comment
  2. File description comment, including purpose of program, inputs, and outputs
  3. `source()` and `library()` statements
  4. Function definitions
  5. Executed statements

## Organization: Layout of .Rmd
- **General Layout:**
We will generally work with R Markdown files (.Rmd) in this class.

- Code comments can be included in R chunks

- Print out all code in documents

- For reproducibility R output can be integrated into text, using "r mean(x)" (using back ticks in place of quotes). _DO NOT_ hard code results in written text.

- Look at output to verify results look how you intended.
  
## Organization: Commenting
>- Comment your code. Entire commented lines should begin with `#` and then one space.
>- Short comments can be placed after code preceded by two spaces, `#` and then one space.
```{r noeval, eval=FALSE}
# create plot of housing price by zipcode
plot(Seattle$Price ~ Seattle$Zip,
  rgb(.5,0,0,.7),  # set transparency for points
  xlab='zipode')
```

## Parting Words
>- Use common sense and *be consistent*.
>- If you are editing code, take a few minutes to look at the code around you and mimic the style.
>- Enough about writing code; the code itself is much more interesting. Have fun!

# Course Goals

With this class, we cannot cover every possible situation that you will encounter. The goals are to:

1. Give you a broad range of tools that can be employed to manipulate, visualize, and analyze data, and
2. teach you to find help when you or your code "gets stuck".


## Process for writing code

When writing code (and conducting statistical analyses) an iterative approach is a good strategy.

1. Test each line of code as you write it and if necessary confirm that nested functions are giving the desired results.
2. Start simple and then add more complexity.

## Debugging Overview

> Finding your bug is a process of confirming the many things that you believe are true -- until you find one which is not true. 
- Norm Matloff

## Debugging Guide 
We will first focus on debugging when an error, or warning is tripped.

1. Realize you have a bug (if error or warning, read the message)
2. Make it repeatable
3. Identify the problematic line (using print statements can be helpful)
4. Fix it and test it (evaluate nested functions if necessary)

## Warnings vs. Errors

R will flag, print out a message, in two cases: warnings and errors.

 - What is the difference between the two?
 - Is the R process treated differently for errors and warnings?
 
## Warnings vs. Errors

- Fatal errors are signaled with `stop()` and force all execution of code to stop triggering an `error`.
- Warnings are generated with `warning()` and display potential problems. Warnings **do not** stop code from executing.
- Messages can also be passed using `message()`, which pass along information.


## Bugs without warning/error

In other cases, we will have bugs in our code that don't necessarily give a warning or an error.

- How do we identify these bugs?
- How can we exit a case where:
    - R is running and may be stuck?
    - the code won't execute because of misaligned parenthesis, braces, brackets?

Note: `NA` values often return a warning message, but not always.


# Tables for R Markdown

## Kable function
Note that output from R can often be hard to read. Luckily there are several options for creating nicely formatted tables. One, which we will use, is the kable function.

## Kable function
```{r kable, mysize=TRUE, size='\\small'}
library(knitr)
kable(aggregate(Loblolly$height,
      by=list(Loblolly$age),mean), digits=3,
      caption= 'Average height of loblolly pine by age',
      col.names = c('Tree Age','Height (ft)'))
```

# Built in R Functions

## R Functions
>- To get more details in R, type `?FunctionName`. This will open up a help window that displays essential characteristics of the function. For example, with the `mean` function the following information is shown:

**Description**: function for the (trimmed) arithmetic mean.

**Usage**: mean(x, trim = 0, na.rm = FALSE, ...)

**x**: An R object.

**trim:** the fraction (0 to 0.5) of observations to be trimmed from each end of x before the mean is computed. 

**na.rm:** a logical value indicating whether NA values should be stripped before the computation proceeds.

## Downloading R Packages
>- R has a set of built in functions, which we have used thus far.
>- R also has a vast repository of "packages" that contain additional, specialized functions. One example is a graphics packaged called `ggplot2` which we will see later in this class.

Using these external packages requires two steps:

1. Download the package `install.packages('ggplot2')`. This only needs to be done once.
2. Load the package `library(ggplot2)`. This needs to be done when opening R.
    
