---
title: "Week 4: Exercises"
output: html_document
---

```{r setup, include=FALSE}
library(knitr)
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
```

## Exercise: Conditions in R 
We have touched on many of these before, but here are some examples of expressions (conditions) in R. Evaluate these expressions:

```{r, eval=FALSE}
pi > 3 & pi < 3.5
c(1,3,5,7) %in% 1:3
1:3 %in% c(1,3,5,7)
rand.uniform <- runif(n = 1, min = 0, max = 1)
rand.uniform < .5
```



## Exercise: Conditional Expression
Write a conditional statement that takes a playing card with two arguments:

- (`card.number <- '4'` and `card.suit <- 'C'` = 4 of clubs or `card.number <-'K'` and `card.suit <- 'H'` = King of hearts)  and
- Print `Yes` if the card is a red face card and `No` otherwise.

Verify this works using the following inputs:

- `card.number <- 'J'` and `card.suit <- 'D'`
- `card.number <- 'Q'` and `card.suit <- 'S'`


## Exercise: Loops

Assume you plan to wager $1 on red for ten roulette spins. If the outcome is red you win a dollar and otherwise you lose a dollar. Write a loop that simulates ten rolls and determines your net profit or loss.

```{r roulette}
RouletteSpin <- function(num.spins){
  # function to simulate roulette spins
  # ARGS: number of spins
  # RETURNS: result of each spin
  outcomes <- data.frame(number = c('00','0',
                as.character(1:36)),
                color=c('green','green','red','black','red','black',
                        'red','black','red','black','red','black',
                        'black','red','black','red','black',
                        'red','black','red','red','black',
                        'red','black','red','black','red',
                        'black','red','black','black','red',
                        'black','red','black','red','black','red'))
  return(outcomes[sample(38,num.spins,replace=T),])
}
```

```{r}
#hint: to get color from a single spin use
RouletteSpin(1)[2]
```



## Exercise: Probability of Red, Green, and Black
1. Calculate/Derive the probability of landing on green, red, and black.

2. How can the `RouletteSpin()` function be used to compute or approximate these probabilities?

## Exercise: Simulation Questions - Part 2 

Now what happens if we:

1. run the simulation again with the same number of trials?

2. run the simulation with more trials, say 1 million?




